---
title: "metExplore: Explore The Met Collection API"
output: rmarkdown::html_vignette
author: "Jiaqi Yao"
vignette: >
  %\VignetteIndexEntry{metExplore: Explore The Met Collection API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```
Purpose / Overview
metExplore is a lightweight R wrapper around The Metropolitan Museum of Art Collection API.
The goal of this package is to make it easy to:
discover available departments in the collection,
search objects with filters (e.g., department, images),
retrieve metadata for one object or many objects,
get analysis-ready results that work well with the tidyverse.
This vignette walks through a complete workflow: departments → search → fetch → batch fetch → simple plot → reproducibility.

To install the metExplore package, please run:

remotes::install_github("ChoWingDev/Data534_Museum_of_Art_API")


Load metExplore plus a couple of tidyverse packages for data manipulation and plotting.
```{r}
library(metExplore)
library(dplyr)
library(ggplot2)

```

A common first step is to see the valid department IDs so we can filter searches to a specific part of the museum.
```{r}
deps <- met_departments()
deps

```

```{r}
dept_id <- deps$department_id[1]
dept_id

```

Search with filters (met_search())
Now we search for objects related to the keyword "cat".
We also request objects that have images, and we restrict to the chosen department.
met_search() returns a tibble of object_id values, and stores the total number of matches as an attribute.
```{r}
res_search <- met_search(
  q = "cat",
  has_images = TRUE,
  department_id = dept_id
)

attr(res_search, "total")
head(res_search, 10)

```

Sometimes a specific department might return zero results depending on the keyword.
If that happens, we fall back to searching across all departments while still requiring images.
```{r}
if (nrow(res_search) == 0) {
  res_search <- met_search(q = "cat", has_images = TRUE)
}

attr(res_search, "total")
head(res_search, 10)


```

Fetch one object (met_object())
Once we have an object_id, we can retrieve detailed metadata for a single object.
```{r}
one_id <- res_search$object_id[1]
one_id

```


```{r}
obj_one <- met_object(one_id)
obj_one

```

Fetch many objects (met_objects())
For analysis and visualization, it is common to retrieve a batch of objects.
Here we take the first 10 IDs from the search results and fetch them together.
```{r}
ids <- head(res_search$object_id, 10)
obj_many <- met_objects(ids)
dplyr::glimpse(obj_many)

```

Output and a simple plot (ggplot2)
As a simple example, we count how many objects in this small sample belong to each department,
then visualize the distribution.
```{r}
plot_df <- obj_many %>%
  filter(!is.na(department)) %>%
  count(department, sort = TRUE)

plot_df

```
```{r}
ggplot(plot_df, aes(x = reorder(department, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    x = "Department",
    y = "Number of objects (sample)",
    title = "Objects by department (sample from search results)"
  )

```

Reproducibility
Finally, we record session details so results can be reproduced.
```{r}
sessionInfo()
```

